---
title: "Clustering companies by consumer complaints"
author: "Douglas Zickuhr"
date: "21/06/2018"
output:
  html_document: default
  pdf_document: default
---

# Loading libraries
```{r libraries, echo=FALSE}
library(tidyverse)
library(e1071)
library(cluster)
library(inaparc)
library(factoextra)

```

# Loading the raw dataset
```{r reading, echo=FALSE}
df <- read_csv('data/dataset.csv')
```

## Checking the dataframe
```{r check data}
head(df)
glimpse(df)
```

## Aggregating the dataset to create indicators by company.
It's important to mention that just companies with more the 50 complaints and at least 5 solved complaints were considered.
```{r df_aggregation}
set.seed(123)
companies <- df %>%
  dplyr::filter(!is.na(company_name)) %>%
  dplyr::mutate(company_name = str_replace_all(company_name,"'","")) %>%
  dplyr::mutate(company_name = str_replace_all(company_name,'"',"")) %>%
  dplyr::mutate(company_name = str_trim(company_name)) %>%
  dplyr::mutate(average_weeks = difftime(complaint_closed_date,
                                  complaint_entered_date,
                                  units="weeks")) %>%
  dplyr::group_by(company_name) %>%
  dplyr::summarise(total_complaint = n(),
            total_solved_complaint = sum(complaint_attended == "Y"),
            perc_solved = total_solved_complaint/total_complaint,
            average_weeks = as.numeric(round(mean(average_weeks),digits = 2))
            ) %>%
  dplyr::filter(total_complaint > 50) %>%
  dplyr::filter(total_solved_complaint > 5) %>%
  dplyr::select(-c("total_solved_complaint")) %>%
  dplyr::mutate(company_id = row_number())
```

## Scalling companies's indicators
```{r scalling}
companies_scalled <- companies %>%
  dplyr::select(-c("company_name","company_id")) %>%
  scale(center = TRUE,
        scale = TRUE)

```


## PCA - Principal component analysis

PCA is a algorithm to apply reduction of dimensions
```{r pca}
companies_pca <- prcomp(companies_scalled, 
                        scale. = FALSE)
```

Looking to the summary of the PCA it's possible to see that with PC2 it's possible to explain more than 70% of data variance
```{r pca_summary}
summary(companies_pca)
```

# Clustering

## Binding the companies data frame and principal components 
```{r companies_pca_binding}
companies_with_pca <- cbind(companies,
                   companies_pca$x)
```


## K-means 
### Running silhouette analysis to find the best number of clustering
```{r kmeans_silhoette_analysis}
companies_silhouette_vector <- numeric(9)
k_var <- 2:10
nstart = 20

for (k in k_var){
  set.seed(123)
  companies_kmeans <- kmeans(companies_with_pca %>%
                               select(starts_with("PC")),
                             centers = k,
                             nstart = nstart,
                             iter.max = 300)
  
  
  companies_kmeans_distance <- companies_with_pca %>%
    select(starts_with("PC")) %>%
    dist()
  
  s <- silhouette(companies_kmeans$cluster,companies_kmeans_distance)
  
  companies_silhouette_vector[k-1] <- mean(s[,3])
}


silhouette_analysis <- tibble(
  clusters = k_var,
  s = companies_silhouette_vector
) %>%
  arrange(desc(s))

silhouette_analysis
```

### Running Kmeans with the best result from silhouette analysis
```{r kmeans}
set.seed(123)

companies_kmeans <- kmeans(companies_with_pca %>%
                  select(starts_with("PC")),
                centers = 5,
                nstart = nstart,
                iter.max = 300)

sil_kmeans <- silhouette(companies_kmeans$cluster,dist(companies_with_pca %>%
              select(starts_with("PC"))))

mean(sil_kmeans[,3])
```


## Fuzzy Clustering
```{r fcm}
set.seed(123)
companies_fcm <- cmeans(x = companies_with_pca %>%
                                 select(starts_with("PC")),
                        centers = 5,
                        iter.max = 500)
```


## Evaluation of values by clusters (Kmeans and Fuzzy Clusters)
### Binding the clusters into a single tibble
It is also joining the FCM membership degree to the dataframe
```{r binding clusters to dataframe}
companies_membership_degree <- companies_with_pca %>%
  cbind(companies_fcm$membership) %>%
  dplyr::rename("fuzzy_cluster_1" = `1`,
                "fuzzy_cluster_2" = `2`,
                "fuzzy_cluster_3" = `3`,
                "fuzzy_cluster_4" = `4`,
                "fuzzy_cluster_5" = `5`)

companies_membership_degree <- companies_membership_degree %>%
  dplyr::select(company_name,starts_with("fuzzy_cluster_")) %>%
  gather(degree,membership_degree,-company_name) %>%
  group_by(company_name) %>%
  dplyr::summarise(fuzzy_membership_degree = max(membership_degree))

companies_clusters <- companies_with_pca %>%
  cbind(kmeans_cluster = companies_kmeans$cluster) %>%
  cbind(fuzzy_cluster = companies_fcm$cluster) %>%
  inner_join(companies_membership_degree)

write_csv(x = companies_clusters, path = "data/clustered_companies.csv")
```

### Average values by Kmeans Clusters
```{r average values for kmeans}
companies_clusters %>%
  dplyr::group_by(kmeans_cluster) %>%
  dplyr::summarise(total_complaint = mean(total_complaint),
                   perc_solved = mean(perc_solved),
                   average_weeks = mean(average_weeks),
                   n = n()) %>%
  dplyr::select(c(kmeans_cluster,total_complaint,perc_solved,average_weeks,n))
```




### Average values by Fuzzy Clusters
```{r average values for hclust}
companies_clusters %>%
  dplyr::group_by(fuzzy_cluster) %>%
  dplyr::summarise(total_complaint = mean(total_complaint),
                   perc_solved = mean(perc_solved),
                   average_weeks = mean(average_weeks),
                   n = n()) %>%
  dplyr::select(c(fuzzy_cluster,total_complaint,perc_solved,average_weeks,n))
```

## Loading the dataset
```{r read sumarised dataset, echo=FALSE}
clustered_companies <- read_csv('data/clustered_companies.csv')
```

## Plots

### Biplot
It is a good feature analyse the data distribution after the dimension reduction
```{r pca biplot}
biplot <- fviz_pca_biplot(X = companies_pca,
                label = "var",
                col.var = "red",
                title = "Consumer Complaints - PC1 and PC2",
                addEllipses  = TRUE,
                ggtheme = theme_bw(),
                alpha = 0.3)

biplot
ggsave(filename = "img/biplot.png",
       plot = biplot)

```

### Elbow Plot
What is the optimal number of clusters for our dataset?
```{r elbow}
return_wss <- function(k, df) {
  x <- kmeans(x = df, centers = k, nstart = 20)
  x$tot.withinss
}

elbow <- tibble(cluster = 2:10, tot.withinss = map_dbl(2:10,return_wss,df = clustered_companies %>%
                                                select(starts_with("PC")))) %>%
  ggplot(aes(x=cluster,y=tot.withinss)) + 
  geom_line() + 
  geom_point() + 
  geom_point(data=tibble(x=4,y=1456.7857), 
             aes(x=x, y=y), colour="red", size=10, alpha = 0.5) +
  labs(title = "Elbow plot",
       x = "Number of clusters k",
       y = "Total within-clusters sum of squares") + 
  theme_bw()

ggsave("img/elbow.png",
       plot = elbow)

elbow
```

### Cluster plot
#### K-means Algorithm result-set
```{r}
kmeans_cluster_viz <- fviz_cluster(companies_kmeans,
             companies %>%
               select(total_complaint,perc_solved,average_weeks) %>%
               scale(),
             geom = "point", 
             shape = 16,
             ggtheme = theme_minimal(),
             alpha = 0.5)

ggsave("img/kmeans_clusters_viz.png",
       plot = kmeans_cluster_viz)

kmeans_cluster_viz
```

#### Fuzzy C-means Algorithm result-set
```{r}
fcm_cluster_viz <- fviz_cluster(companies_fcm,
             companies %>%
               select(total_complaint, perc_solved, average_weeks),
             geom = "point", 
             shape = 16,
             ggtheme = theme_minimal(),
             alpha = companies_clusters$fuzzy_membership_degree)

ggsave("img/fcm_clusters_viz.png",
       plot = fcm_cluster_viz)

fcm_cluster_viz
```

### Silhouette plot
#### K-Means Algorithm result-set
```{r}
kmeans_dist <- dist(scale(clustered_companies %>% select(starts_with("PC"))))
sil <- silhouette(clustered_companies$kmeans_cluster,kmeans_dist)

silplot <- fviz_silhouette(sil, 
                           label = FALSE, 
                           print.summary = TRUE,
                           scale_fill_brewer = scale_fill_brewer(palette = "Set2"))
ggsave(filename = "img/kmeans_silhouette.png",
       plot = silplot)

silplot
```

#### Fuzzy C-Means Algorithm result-set
```{r}
fcm_dist <- dist(scale(clustered_companies %>% select(starts_with("PC"))))
fcm_sil <- silhouette(clustered_companies$fuzzy_cluster,fcm_dist)

fcm_silplot <- fviz_silhouette(fcm_sil, 
                           label = FALSE, 
                           print.summary = TRUE,
                           scale_fill_brewer = scale_fill_brewer(palette = "Set2"))

fcm_silplot

ggsave(filename = "img/fcm_silhouette.png",
       plot = fcm_silplot)
```

### Clusters Features
#### Size of the clusters

```{r}
clusters_size <- tibble(algorithm = "K-means",cluster = 1:5 ,size = companies_kmeans$size) %>%
  bind_rows(tibble(algorithm = "Fuzzy C-means", cluster = 1:5, size = companies_fcm$size)) %>%
  group_by(algorithm) %>%
  dplyr::mutate(perc = size/sum(size)*100,
         total = sum(size),
         no = n()) %>%
  ungroup() %>%
  dplyr::mutate(cluster = reorder(cluster,desc(cluster))) %>%
  dplyr::mutate(algorithm = factor(algorithm),
                algorithm = fct_rev(algorithm)) %>%
  ggplot(aes(x = cluster, y=size+10, 
             fill = algorithm),
         colour = "black") + 
  geom_bar(stat = "identity",
           show.legend = FALSE) + 
  geom_text(aes(y = size + 60, 
               label = paste0(round(perc,2),"%")
               )) + 
  facet_wrap(~algorithm, scales = "fixed" , ncol = 1) + 
  scale_fill_brewer(palette = "Set2") + 
  labs(title = "Size of Clusters by Algorithm",
       x = "Cluster",
       y = "Number of Observations") + 
  theme_minimal() + 
  coord_flip()

clusters_size

ggsave("img/clusters_size.png",clusters_size)
```

#### Distribution of variables - K-Means
```{r}
kmeans_dist_plot <- companies_clusters %>%
  select(total_complaint,perc_solved,average_weeks,kmeans_cluster) %>%
  gather(feature,values,total_complaint:average_weeks) %>%
  dplyr::mutate(feature = factor(feature),
                feature = fct_recode(feature,
                                     `Total Number of Complaints` = "total_complaint",
                                     `Solving Rate (%)` = "perc_solved",
                                     `Average Duration (Weeks)` = "average_weeks")) %>%
  dplyr::mutate(values = if_else(feature == "Solving Rate (%)",values * 100, values)) %>%
  ggplot() + 
  geom_jitter(aes(x=kmeans_cluster,
                  y=values,
                  colour=factor(kmeans_cluster)), 
              alpha = 0.1) +
  geom_boxplot(aes(x=kmeans_cluster,
                   y=values,
                   group=kmeans_cluster,
                   fill=factor(kmeans_cluster)),
               outlier.shape = NA) + 
  ggplot2::scale_y_log10() + 
  facet_wrap(~feature,
             scales = "free") + 
  labs(title = "Distribution of values by Feature by Cluster",
       x = "Cluster",
       y = "Values",
       fill = "Cluster") +
  theme_minimal() + 
  scale_fill_brewer(palette = "Set2") + 
  scale_color_brewer(palette = "Set2") + 
  guides(colour=FALSE)

ggsave(filename = "img/kmeans_distribution.png",
       plot = kmeans_dist_plot,
       width = 12,
       height = 8)

kmeans_dist_plot
```

#### Distribution of variables - Fuzzy C-Means
```{r}
fcm_dist_plot <- companies_clusters %>%
  select(total_complaint,perc_solved,average_weeks,fuzzy_cluster) %>%
  gather(feature,values,total_complaint:average_weeks) %>%
  dplyr::mutate(feature = factor(feature),
                feature = fct_recode(feature,
                                     `Total Number of Complaints` = "total_complaint",
                                     `Solving Rate (%)` = "perc_solved",
                                     `Average Duration (Weeks)` = "average_weeks")) %>%
  dplyr::mutate(values = if_else(feature == "Solving Rate (%)",values * 100, values)) %>%
  ggplot() + 
  geom_jitter(aes(x=fuzzy_cluster,
                  y=values,
                  colour=factor(fuzzy_cluster)), 
              alpha = 0.1) +
  geom_boxplot(aes(x=fuzzy_cluster,
                   y=values,
                   group=fuzzy_cluster,
                   fill=factor(fuzzy_cluster)),
               outlier.shape = NA) + 
  ggplot2::scale_y_log10() + 
  facet_wrap(~feature,
             scales = "free") + 
  labs(title = "Distribution of values by Feature by Soft Cluster",
       x = "Cluster",
       y = "Values",
       fill = "Cluster") +
  theme_minimal() + 
  scale_fill_brewer(palette = "Set2") + 
  scale_color_brewer(palette = "Set2") + 
  guides(colour=FALSE)

ggsave(filename = "img/fcm_distribution.png",
       plot = fcm_dist_plot,
       width = 12,
       height = 8)

fcm_dist_plot
```
